Evaluation of Types

### 2020-05-15 evaluation/R/evaluate-type-analysis.Rmd

---
title: Evaluation of Types
output:
  html_document:
    theme: cerulean
    toc: true
    toc_float: true
params:
  type_analysis_datafile: ../data/type-analysis/merged/merged.csv
  graph_dirpath: ../../paper/plots/
---

@@

suppressPackageStartupMessages(library(stringr))
suppressPackageStartupMessages(library(dplyr))
suppressPackageStartupMessages(library(readr))
suppressPackageStartupMessages(library(purrr))
suppressPackageStartupMessages(library(tidyr))
suppressPackageStartupMessages(library(DT))
suppressPackageStartupMessages(library(ggplot2))
suppressPackageStartupMessages(library(tibble))

@@

@@

infer_kind <- function(type) {
    if(type %in% c("integer", "double", "character", "logical", "complex", "raw")) {
        "scalar"
    }
    else if(type %in% c("^integer", "^double", "^character", "^logical", "^complex")) {
        "scalar-na"
    }
    else if(type %in% c("integer[]", "double[]", "character[]", "logical[]", "complex[]", "raw[]")) {
        "vector"
    }
    else if(type %in% c("^integer[]", "^double[]", "^character[]", "^logical[]", "^complex[]")) {
        "vector-na"
    }
    else if(str_detect(type, "^list")) {
        "list"
    }
    else if(str_detect(type, "^struct")) {
        "struct"
    }
    else if(str_detect(type, "^class")) {
        "class"
    }
    else if(type == "any") {
        "any"
    }
    else if(type == "null") {
        "null"
    }
    else if(type == "...") {
        "..."
    }
    else {
        "other"
    }
}

type_analysis_data <-
   type_analysis_data %>%
   mutate(kind = map_chr(type, infer_kind))

@@

union_types <-
union_data %>%
group_by(package_name, function_name, parameter_position, alternative_count) %>%
summarize(type = paste0(sort(type), collapse=" | ")) %>%
ungroup() %>%
group_by(alternative_count, type) %>%
summarize(count = n()) %>%
ungroup() %>%
group_by(alternative_count) %>%
arrange(desc(count)) %>%
ungroup() %>%
group_by(alternative_count) %>%
mutate(percentage = round(100 * count / sum(count), 2),
       cumsum_percentage = round(100 * cumsum(count) / sum(count), 2)) %>%
ungroup() %>%
group_by(alternative_count) %>%
slice(1:20) %>%
ungroup()

@@

>>>>

---
title: Evaluation of Types
output:
  html_document:
    theme: cerulean
    toc: true
    toc_float: true
params:
  type_analysis_datafile: ../data/type-analysis/merged/merged.csv
  graph_dirpath: ../../paper/plots/
  table_dirpath: ../../paper/tables/
---

@@

suppressPackageStartupMessages(library(stringr))
suppressPackageStartupMessages(library(dplyr))
suppressPackageStartupMessages(library(readr))
suppressPackageStartupMessages(library(purrr))
suppressPackageStartupMessages(library(tidyr))
suppressPackageStartupMessages(library(DT))
suppressPackageStartupMessages(library(ggplot2))
suppressPackageStartupMessages(library(tibble))
suppressPackageStartupMessages(library(knitr))
suppressPackageStartupMessages(library(kableExtra))

@@

save_table <- function(content, filename) {
    write(content, file = file.path(params$table_dirpath, filename))
}

simplify_types <- function(type) {
    if(type %in% c("any", "externalptr", "expression", "pairlist")) {
        "any"
    }
    else {
        type
    }
}

is_polymorphic <- function(kind) {
    kind %in% c("...", "any", "list")
}

infer_kind <- function(type) {
    if(type %in% c("integer", "double", "character", "logical", "complex", "raw")) {
        "scalar"
    }
    else if(type %in% c("integer[]", "double[]", "character[]", "logical[]", "complex[]", "raw[]")) {
        "vector"
    }
    else if(type %in% c("^integer[]", "^double[]", "^character[]", "^logical[]", "^complex[]", 
                        "^integer", "^double", "^character", "^logical", "^complex")) {
        "^vector"
    }
    else if(str_detect(type, "^list")) {
        "list"
    }
    else if(str_detect(type, "^struct")) {
        "struct"
    }
    else if(str_detect(type, "^class")) {
        "class"
    }
    else if(type == "any") {
        "any"
    }
    else if(type == "null") {
        "null"
    }
    else if(type == "...") {
        "..."
    }
    else if(type == "environment") {
        "environment"
    }
    else {
        "other"
    }
}

@@

type_analysis_data <-
    type_analysis_data %>%
    mutate(type = map_chr(type, simplify_types)) %>%
    mutate(kind = map_chr(type, infer_kind))

@@

union_types <-
union_data %>%
group_by(package_name, function_name, parameter_position, alternative_count) %>%
summarize(kind = paste0(sort(kind), collapse=" | ")) %>%
ungroup() %>%
group_by(alternative_count, kind) %>%
summarize(count = n()) %>%
ungroup() %>%
group_by(alternative_count) %>%
arrange(desc(count)) %>%
ungroup() %>%
group_by(alternative_count) %>%
mutate(percentage = round(100 * count / sum(count), 2),
       cumsum_percentage = round(100 * cumsum(count) / sum(count), 2)) %>%
ungroup() %>%
group_by(alternative_count) %>%
slice(1:20) %>%
ungroup()

@@

unityped_positions <-
union_types %>% 
filter(alternative_count == 1)

polymorphic_unityped_rows <-
unityped_positions %>%
pull(kind) %>% 
map_lgl(is_polymorphic)

polymorphic_unityped_rows <- (1:nrow(unityped_positions))[polymorphic_unityped_rows]

print(polymorphic_unityped_rows)

unityped_positions %>%
select("Types" = kind, "Parameter #" = count, "%" = percentage, "Cumulative %" = cumsum_percentage) %>%
kable("latex", booktabs = T) %>%
kable_styling(full_width = F, latex_options = c("scale_down")) %>%
row_spec(polymorphic_unityped_rows, color = "black", background = "lightgray") %>%
save_table("uni-typed-breakdown.tex")
